Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-08-21 draft.tex
Rewrote conclusion to sum up the main rationales.

% 1. parameters and 2. code version with results
This is why results files should come alongside their context, i.e. an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does. The code itself is part of that context: the version of the code must be recorded as part of the provenance data. It is common for different results or different figures to have been generated by different versions of the code. Ideally, all results should originate from the same (and last) version of the code. But for long or expensive computations, this may not be feasible. In that case, the result files should contain the version of the code that was used to produce it. This information can be obtained from the version control software. This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed. In R$^3$, the code records the last git commit hash, and whether the repository holds uncommited changes when the computation starts.\\

% scripts to generates figures & data behind graphs
Published results should come from version of the code where every file has been commited. This includes pre-processing, post-processing and plotting code. Plotting code may seem mundane, but it is as vulnerable as any other piece of the code to bugs and errors. With a code that is re-runnable, repeatable and contains plotting scripts, a user can quickly reproduce published figures. When it comes to checking that the reproduced data match the one published in the article, however, figures can reveal themselves to be imprecise and cumbersome, and sometimes plain unusable. Ideally, published figures should be accompanied by their underlying data (coordinates of the plotted points) in the supplementary data to allow straightforward numeric comparisons.\\

@@

def git_info():
    """Retrieve the current git hash, and if the repository has uncommited changes."""
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
            'git_version': str(pip.git.Git().get_git_version())}

@@

Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With a documention explaining which script or configuration file correspond to which experiment, reproducing the different figures becomes straightforward.\\

@@

%FIXME: I feel like this paragraph is openning me to attack in a way that is probably unecessary.
Striving, as we do here, for a perfect quantitative match may seem unnecessary. Yet, in replication projects, quantitative comparisons are an effective way to verify that the behavior has been reproduced. Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.\\

Here we are able to replicate exactly the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\

But there are still more subtle problems with the description given above.
If we look more closely at it, we can realize
that nothing is said about the specific case of {\tt 0} when generating a step.
Do we have to consider {\tt 0} to be a positive or a negative step? Without
further information and without the original code, it is up to the reader to
decide. Likewise, the description is ambiguous regarding the first element of the walk. Is the initialization value included (it was not in our codes so far)? This slight difference affects statistics, especially for short runs.\\

There is no better way to ferret out the ambiguities, big and small, of an
article than to replicate it. This is one of the reasons why the ReScience
journal \citep{Rougier:2017} has been created. This journal {\em targets
  computational research and encourages the explicit replication of already
  published research, promoting new and open-source implementations in order to
  ensure that the original research is reproducible}.\\

@@

@@

%FIXME: can still be improved
Through a small random walk example implemented in Python, we've seen that writing good scientific code is not trivial. The
R$^0$ form, while easy to write and simple to understand is not a good
scientific code. Adding a single comment, we transformed this code into the
R$^1$ version (rerunnable) that partially solved the common dependencies
problem and will allow us to re-run the program two months or two years from
now. In the meantime, the R$^2$ version (repeatable) guarantees us (to some
extent) we can faithfully repeat our experiment from one run to the other.  Of
course, the full-fledge R$^4$ (reusable) form is much preferred but it requires
a fair amount of work that might be too much of a burden. R$^3$ (reproducible)
form seems to be an acceptable compromise and should be accepted as the minimum
scientific standard \citep{Wilson:2017}. This means this should be actually
checked by reviewers and publishers when code is part of a work worth being
published. But it's hardly the case today.\\

>>>>

% 1. parameters and 2. code version with results
This is why results files should come alongside their context, i.e. an exhaustive list of the parameters used as well as a precise description of the execution environment, as the R$^3$ code does. The code itself is part of that context: the version of the code must be recorded as part of the provenance data. It is common for different results or different figures to have been generated by different versions of the code. Ideally, all results should originate from the same (and last) version of the code. But for long or expensive computations, this may not be feasible. In that case, the result files should contain the version of the code that was used to produce it. This information can be obtained from the version control software. This also allows, if some errors are found and corrected after some results have been obtained, to identify which ones should be recomputed. In R$^3$, the code records the last git commit hash, and whether the repository holds uncommitted changes when the computation starts.\\

% scripts to generates figures & data behind graphs
Published results should come from version of the code where every file has been committed. This includes pre-processing, post-processing and plotting code. Plotting code may seem mundane, but it is as vulnerable as any other piece of the code to bugs and errors. With a code that is re-runnable, repeatable and contains plotting scripts, a user can quickly reproduce published figures. When it comes to checking that the reproduced data match the one published in the article, however, figures can reveal themselves to be imprecise and cumbersome, and sometimes plain unusable. Ideally, published figures should be accompanied by their underlying data (coordinates of the plotted points) in the supplementary data to allow straightforward numeric comparisons.\\

@@

def git_info():
    """Retrieve the current git hash, and if the repository has uncommitted changes."""
    repo = git.Repo(os.path.dirname(__file__), search_parent_directories=True)
    return {'hash': repo.head.object.hexsha, 'dirty': repo.is_dirty(),
            'git_version': str(pip.git.Git().get_git_version())}

@@

Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be explicitly tracked in the result files since they contain the parameters used, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\

@@

Here we are able to replicate exactly\footnote{Striving, as we do here, for a perfect quantitative match may seem unnecessary. Yet, in replication projects, quantitative comparisons are an effective way to verify that the behavior has been reproduced. Moreover, they are particularly helpful to track exactly where the code of a tentative replication fails to reproduce the published results.} the behavior of the pure-Python random walker by setting the internal state of the NumPy RNG appropriately, but only because we have access to specific technical details (the use of the {\tt random} module of the standard Python library of CPython 3.6.1), or to the code itself.\\

@@

But there are still more subtle problems with the description given above.
If we look more closely at it, we can realize
that nothing is said about the specific case of {\tt 0} when generating a step.
Do we have to consider {\tt 0} to be a positive or a negative step? Without
further information and without the original code, it is up to the reader to
decide. Likewise, the description is ambiguous regarding the first element of the walk. Is the initialization value included (it was not in our codes so far)? This slight difference affects the statistics of short runs.\\

All these ambiguities in the description of an algorithm pile up; some are inconsequential (the 0 case has null probability), but some may affect the results in important ways. They are mostly inconspicuous to the reader and oftentimes, to the writer as well. In fact,  there is no better way to ferret out the ambiguities, big and small, of an article than to replicate it. This is one of the reasons why the ReScience journal \citep{Rougier:2017} has been created. This journal {\em targets  computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\

@@

Throughout the evolution of a small random walk example implemented in Python, we illustrated some of the issue that may plague code that work and produce correct results, and reduce its value as an element of scientific knowledge.
Further, we articulated characteristics that a code should possess to be a useful part of a scientific publication: it should be rerunnable, repeatable, reproducible, reusable and replicable.\\

Running old code on tomorrow’s computer and software stacks may not be possible. But recreating the old code’s execution environment may be: to ensure that the long-term rerunnability of a code, its execution environment must be documented. For our example, a single comment went a long way to go from the R$^0$ to the R$^1$ (rerunnable) code.\\

Science is built on verifying the results of others. This is harder to do if each execution of the code produce a different result. While for complex parallel workflow this may not be possible, in all instances where it is feasible the code should be repeatable. This allows future researchers to examine exactly how a specific result was produced. Most of the time, what is needed is to set or record the initial state of the pseudo-random number generator, as what done in the R$^2$ (repeatable) version.\\

Even more care is needed to make a code reproducible. The exact execution environment, code and parameters used must be recorded and embedded in the results files. The R$^3$ (reproducible) version implements provenance methods to that end. Furthermore, the code must be made available as supplementary data with the whole computational workflow, from preprocessing steps to plotting scripts.\\

Making code reusable is a stretch goal that can yield tremendous benefits for you, your team and other researchers. Taken into account during development rather than as an afterthought, simple measures can avoid hours of head-scratching for others, and for yourself—in a few years. Documentation is paramount here, even a single comment per function, as was done in the R$^4$ (reusable) version.\\

Finally, there is the belief that an article should suffice itself: the description of the algorithms present in the paper should suffice to re-obtain (to replicate) the published results. For well-written papers that precisely dissociate conceptually significant details from irrelevant implementation details, that may be. But science should not assume the best of cases. Science assumes that errors are ubiquitous. Every paper is a mistake or a forgotten parameter away from irreproducibility. Replication efforts use the paper first, and then the reproducible code that comes along with it whenever the paper falls short of being precise enough to be reimplemented.\\

In conclusion, the R$^3$ (reproducible) form should be accepted as the minimum
scientific standard \citep{Wilson:2017}. This means this should be actually
checked by reviewers and publishers when code is part of a work worth being
published. But it's hardly the case today.\\
